Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems\thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}

### 2020-10-02 2020/OPTIMA/optima_paper.tex
OPTIMA: fixes, update slides
OPTIMA: add revised versions of the answers and paper

\begin{document}
%
\title{Parallel global optimization algorithm with uniform convergence
for solving a set of constrained global optimization problems\thanks{This study was supported
by the Russian Science Foundation, project No.\,16-11-10150.}}
%
\titlerunning{Parallel global optimization algorithm with uniform convergence}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov\and
Konstantin Barkalov
}
%
\authorrunning{V. Sovrasov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{sovrasov.vladislav@itmm.unn.ru},
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In the present work, an extension of global optimization algorithm for solving a set of problems
simultaneously onto a case of problems with non-convex constraints is considered. The property
of uniform convergence inherent to the algorithm allows distributing the computation resources
in the optimal way since the errors of numerical solutions decrease with approximately equal
rates in all problems in the course of the execution of the algorithm. The algorithm assigns a
priority to each problem and performs the computations of the objective functions in several
problems in parallel at every iteration. Upon stop of the method at arbitrary time moment,
solutions of similar quality would be obtained in all problems of the series being solved. A series
of several problems arises if the global optimization problem has a discrete parameter or, for
example, in solving a multicriteria optimization problem by applying scalarization techniques. The
considered algorithm employs Peano-type space-filling curves for the reduction of the
multidimensional optimization problems to the one-dimensional ones.  The efficiency of the
implemented algorithm was tested using the sets of artificially generated global optimization
problems as well as in solving a series of problems obtained as a result of scalarization of a
multicriteria optimization  problem. Also, the numerical experiments confirmed the uniform
convergence of the proposed method.

@@

\section{Introduction}
\label{sec-intro}
Global optimization problems with non-convex constraints are considered to be one
of the most difficult optimization problems.
Finding of the global minimum of a function of several variables appears to be more
difficult often than a local optimization in a thousand-dimensional space.
For the latter, application of the simplest gradient descent method may appear to be enough,
whereas to \textit{guarantee} the finding of the global optimum the optimization
methods have to accumulate the information on the behaviour of the objective function in the
whole search domain \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
The solving of a series of such problems at limited computational resources is even more
difficult problem: besides the search of the global extremum, it is necessary to distribute the
computational resources in such a way that the location of the global extremum would be
estimated with approximately equal quality in all problems being solved at once.
Usually, a series of \(q\) problems is solved either sequentially or in parallel by portions of \(p\ll
q\) problems where \(p\) is a number of parallel computational devices.
Such an approach leads to a situation when at every time moment until the stop of computations
some problems remain, in which the estimate of the global optimum is not obtained yet whereas
in the problems from the beginning of the list the optimum may be estimated even with excess
precision.

In the present work, a generalization of parallel global optimization method for simultaneous
solving of a set of problems developed earlier in Lobachevsky State University of Nizhni Novgorod
\cite{BarkalovStrongin2018} onto the case of non-convex constraints is considered. To account
for the constraints, the index scheme \cite{Strongin2000,Pugliese} is applied allowing working with
partially defined objective function and constraints and featured by efficiency comparable to other
approaches \cite{BarkalovLebedev2017}. The convergence of the algorithm to the global
optimizers in all problems was proved theoretically, the efficiency of the implemented
algorithm was demonstrated by solving sets of problems generated by a special mechanism
generating the sets of problems of predefined dimensionality with predefined number of
non-convex constraints \cite{GergelBarkalov2019}.
Besides the problems generated artificially, the considered method was tested also on a set of
problems arising when solving a multicriteria optimization problem with nonlinear constraints by
criteria convolution method \cite{Ehrgott2005}.

The paper is structured as follows. In Section \ref{sec:problem}, the statement of the problem to
be solved is considered. Then, in Section \ref{sec:method}, a description of the parallel
optimization method is given. The sufficient conditions of convergence of the considered
method are given in Section \ref{sec:conv_method}.
The results of the numerical experiments confirming the efficiency of considered method are
presented in Section \ref{sec:exps}.
In Conclusion, a short summary of the results obtained in the present study is given and the
directions of further effort on the improvement of the software implementation of the
considered method are outlined.

\section{Global optimization problem statement}
\label{sec:problem}
In this work we consider the following
problem: to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in
a hyperinterval $D$
\begin{equation}
\label{eq:task}
\varphi(y^*)=\min\{\varphi(y):y\in D\},
\end{equation}
\begin{displaymath}
D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}.
\end{displaymath}
In order to construct an estimate of the global minimum based on a finite number of
computations of the function values, the rate of variation of \(\varphi(y)\) in \(D\) is required to
be limited.
As a rule, the Lipschitz condition
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
\end{displaymath}
is accepted as such a limitation.

There are various methods solving the considered multidimensional problem directly
\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
In the present work, a one-dimensional method is considered, which is applied together with the
dimensionality reduction scheme.
The use of space-filling curve (or \textit{evolvent}) $y(x)$, where
\begin{equation}
\label{cube}
\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\},
\end{equation}
 is the well-known scheme of dimensionality reduction
 of the initial problem for the global optimization algorithms
\cite{Sergeyev2013}.
A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the
one-dimensional function \(\varphi(y(x))\) is not lipschitzian but h\"{o}lderian one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, \; x_1,x_2\in[0,1],
\end{displaymath}
where the H\"{o}lder constant \(H\) is related to the Lipschitz one \(L\) by the relation
\begin{displaymath}
  H=2L\sqrt{N+3}.
\end{displaymath}

@@

Let us denote \(g_{m+1}(y)=\varphi(y)\). Hereafter, we will assume all functions
\(g_k(y),1\leqslant k \leqslant m+1\),
to satisfy the Lipschitz condition in some hyperinterval including \(D\).

@@

@@

Taking into account the dimensionality reduction scheme (\ref{cube}), let us assume when
describing the method that it is required to find the global minimum of the function \(\varphi(x),
x\in[0,1]\), satisfying the H\"{o}lder condition with the constraints \(g_j(x)\) also satisfying this
condition in the interval \([0,1]\).

The index algorithm of global search (IAGS) for solving the one-dimensional problems
%(\ref{eq:constrained_problem})
considered here implies the constructing of a sequence of points
\(x_k\), which the values of index function \(z_k\) are computed in.
On account of, the index scheme \cite{Strongin2000} was used.
Let's assume that each function \(g_i(x), 1 \leqslant i \leqslant m + 1\), is defined and computable
only in the corresponding subrange \(Q_i \in [a, b]\), where \(\varphi(x)\) is denoted as \(g_{m+1}(x)\) and
\begin{equation}
  \label{eq:q}
  Q_1=[0,1],\: Q_{i+1}=\left\{x\in Q_i:g_{i}(x)\leqslant 0\right\},\:1 \leqslant i \leqslant m, \: Q_{m+2}=\emptyset
\end{equation}
In the view of the definition (\ref{eq:q}) the initial problem can be rewritten in the form
\begin{displaymath}
  \varphi(x^*)=\min\{g_{m+1}(x):x\in Q_{m+1}\}.
\end{displaymath}

The index scheme assigns a number called index \(\nu(x)\) to the each point of the search sequence \(x_k\):
\begin{equation}
  \nu(x) = i:x\in Q_i, x\not\in Q_{i+1},\:1\leqslant i\leqslant m + 1.
\end{equation}

In order to obtain an index of the point \(x\) we have to perform a trial defined via the following steps:
\begin{enumerate}
  \item Consequently compute functions \(g_i(x),1 \leqslant i \leqslant m\) until for some \(i^*\) \(g_{i^*}(x)>0\) or until \(i=m+1\).
  Set \(\nu(x)=i^*\) or \(\nu(x)=m+1\) if \(g_i(x) \leqslant 0, 1 \leqslant i \leqslant m\).
  Therefore, \(\nu(x)\) is the number of the first violated constraint at the point \(x\) or \(m+1\) if
  all constraints are satisfied at \(x\).
  \item Return pair \(\nu(x), z=g_{\nu(x)}(x)\) as a result of trial.
\end{enumerate}

Such an approach to conducting the trials allows reducing the initial problem with functional
constraints to an unconstrained problem of minimization of a discontinuous function:
\begin{displaymath}
  \begin{array}{lr}
    \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M,\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
=M.\end{cases}}
  \end{array}
\end{displaymath}
Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
_{}^{}\left\{g_{M}(x):x\in Q_{M}\right\}\).
Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
The function \(\psi (x)\) satisfies the H\"{o}lder condition on the set
\(Q_1\) with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
of the sets \(Q_i,\:1\leqslant i \leqslant m+1\).
Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
known in advance, these ones can be estimated in the course of solving the problem.
A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
accumulated by the method upon executing \(n\) steps.

At the first iteration of the method, the trial is performed in the arbitrary internal point \(x_1\) of
the interval \([0,1]\).
The indices of the points 0 and 1 are considered to be zero ones, the values \(z\) in these points
are undefined.
Let \(k\geqslant 1\) iterations of the method have been performed, in the course of which the
trials were conducted in \(k\) points \(x_i, 1\leqslant i\leqslant k\).
Then, the points \(x^{k+1}\) of the search trials of the next \((k+1)^{\rm th}\) iteration are
defined according to the rules:

Step 1. Renumber the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials by
the lower indices in the order of increasing coordinate values, i.e.
\begin{equation}
  \label{eq:points}
  0=x_0<x_1<\dotsc<x_{k+1}=1,
\end{equation}
and juxtapose to them the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) computed at these points.

Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\), determine the corresponding
set \(I_{\nu }\) of the lower indices of the points, which the values of the functions \(g_{\nu
}(x)\) are computed in:
\begin{displaymath}
  I_{\nu }=\{i:\nu (x_{i})=\nu ,1\leqslant i\leqslant k\},1\leq \nu \leqslant m+1,
\end{displaymath}
and determine the maximum value of the index \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).

Step 3. Compute current estimate for the unknown H\"{o}lder constant:
\begin{equation}
  \label{step2}
  \mu _{\nu }=\max \left\{ \frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-
x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j \right\}.
\end{equation}
If the set \(I_{\nu }\) contains less than two elements or if the value \(\mu _{\nu }\) appears to
be equal to zero, then accept \(\mu _{\nu }=1\).

Step 4. For all nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
\begin{equation}
  \label{eq:step_4}
  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
  0,&\nu <M.\end{cases}}
\end{equation}

Step 5. For each interval \((x_{i-1},x_{i}),1\leqslant i\leqslant k\) compute the characteristic
\begin{equation}
  \label{step3_1}
  R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu
})^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}},&\nu =\nu
(x_{i})=\nu (x_{i-1}),\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu
}}},&\nu =\nu (x_{i-1})>\nu (x_{i}),\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu
_{\nu }}},&\nu =\nu (x_{i})>\nu (x_{i-1}),\end{cases}}
\end{equation}
where \(\Delta_{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\).
The values \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
The products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
unknown H\"{o}lder constants depend on these ones.

@@

The algorithm stops if the condition \(\Delta_{t}\leqslant \varepsilon\) is satisfied where
\(\varepsilon>0\) is a predefined precision.
The values
\begin{equation}
\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), \; x_k^*=\arg \min_{1\leqslant i \leqslant
k}\varphi(x_i)
\end{equation}
are accepted as the estimates of the global solution.

Next, following the approach described in \cite{BarkalovStrongin2018}, we will use \(q\) copies
of IAGS working synchronously to solve the problem series (\ref{eq:many_problems}).
The only difference is that when selecting the interval with the maximum characteristic at Step
6, the choice will be made from all intervals, which \(q\) copies of IAGS have generated to the
moment.
If the maximum characteristic corresponds to the \(i^{\rm th}\) problem, then Step 7 is
performed in the \(i^{\rm th}\) copy of the method while the rest copies stay idle.
This way, at every iteration the trial is performed in the most promising problem from the
viewpoint of the characteristics (\ref{step3_1}) that allows distributing the resources of the
method among the problems dynamically. Let's denote this method as MIAGS.

The parallel modification of the method doesn't differ from the one considered in
\cite{BarkalovStrongin2018} and consists in the selection of \(p\) intervals at Step 6 and the
performing of \(p\) trials in parallel at the next step.
At that, all resources of the method within the framework of the iteration may be directed onto
a single problem as well as onto \(l\leqslant p\) problems simultaneously (subject to the problem,
which the intervals selected by the method belong to).

@@

The conditions of convergence of the described in Section \ref{sec:method} method in case of \(q=1\) are given in \cite{Strongin2000}.
\begin{theorem} (Sufficient convergence conditions)
  \label{th:single_conv}
  Assume that the following conditions are true:
  \begin{enumerate}
    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
  by the index method while solving the problerm
  (\ref{eq:constrained_problem}) is feasible and satisfies the conditions
\begin{equation}
  \label{eq:conv_cond}
  \varphi(\overline{y})=\inf\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant i\leqslant m, k=1,2,\dots\}=\varphi(y^*).
\end{equation}
\end{theorem}

@@

\subsection{Example of solving a multicriteria problems}

>>>>

\begin{document}
%
\title{Parallel global optimization algorithm with uniform convergence
for solving a set of constrained global optimization problems\thanks{This study was supported
by the Russian Science Foundation, project No.\,16-11-10150.}}
%
\titlerunning{Parallel global optimization algorithm with uniform convergence}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov\and
Konstantin Barkalov
}
%
\authorrunning{V. Sovrasov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia
\email{sovrasov.vladislav@itmm.unn.ru},
\email{konstantin.barkalov@itmm.unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This paper proposes an extension of global optimization algorithm for solving a set of problems
with non-convex constraints. The uniform convergence inherent to the algorithm allows for the optimal
distribution of the computational resources, since the number of errors in numerical solutions
decreases at approximately equal rate for all problems processed by the algorithm. The algorithm assigns
a priority level to each problem and, with every iteration, performs the computations of the objective
functions for several problems in parallel. After the algorithm is terminated at any given time,
it obtains solutions of similar quality for all the problems of the set. Such sets appear when
the global optimization problem has a discrete parameter or, for example, when a multicriteria optimization
problem is solved by scalarization techniques. The algorithm employs Peano-type space-filling curves
for the reduction of the multidimensional optimization problems to the one-dimensional ones.
The efficiency of the implemented algorithm was tested using the sets of artificially generated
global optimization problems, as well as a series of problems obtained as a result of scalarization
of a multicriteria optimization problem. Additional numerical experiments also confirmed
the uniform convergence of the proposed method.

@@

\section{Introduction}
\label{sec-intro}
Global optimization problems with non-convex constraints are considered as one of the most difficult
optimization problems.
Finding of the global minimum of a function with several variables often appears to be
more difficult than a local optimization in a thousand-dimensional space.
The simplest gradient descent method could be sufficient for the latter, however,
to \textit{guarantee} that the global optimum is found, the optimization
methods have to accumulate the information on the behavior of the objective function
in the entire search domain  \cite{Jones2009,Paulavicius2011,Evtushenko2013,Strongin2000}.
In some cases, for instance, for DC or QP problems, more effective methods and even
analytical optimality conditions could be applied  \cite{PhamDinh2014, qpBook},
but these methods are barely applicable for black box functions (i.e., an analytical
representation of the functions is not available).

Solving a series of such problems with limited computational resources is even more
difficult: besides the search of the global extremum, the computational resources
have to be distributed in such a way that the location of the global extremum would be
estimated with approximately equal quality for all problems being solved at the same time.
Usually, a series of  \(q\) problems is solved either sequentially or in parallel by sets of  \(p\ll
q\) problems where \(p\) is the number of parallel computational devices.
Such an approach creates a situation when at every given moment there exist problems for which
the global optimum is not obtained yet, while the optimum for the problems from the beginning of
the set may be estimated even with excess precision.

This paper considers a parallel global optimization method developed at the Lobachevsky
State University of Nizhny Novgorod for simultaneous solving of a set of problems \cite{BarkalovStrongin2018}
and in particular, its generalization to cover the non-convex constraints. The index scheme \cite{Strongin2000,Pugliese}
which is applied to account for the constraints, allows operating with partially defined objective functions and constraints, and its efficiency is on par with similar approaches \cite{BarkalovLebedev2017}.
The convergence of the algorithm to the global optimizers in all problems was proved theoretically,
the efficiency of the implemented algorithm was demonstrated on sets of problems generated
by a special mechanism. This mechanism generates sets of problems with predefined dimensionality
and predefined number of non-convex constraints  \cite{GergelBarkalov2019}.
Besides the artificially generated problems, the method was also tested on sets of
multicriteria optimization prob¬lems with nonlinear constraints solved by
criteria convolution method  \cite{Ehrgott2005}.

The paper is structured as follows. Section \ref{sec:problem} provides the statement of the problem.
Section \ref{sec:method} gives a description of the parallel optimization method.
Subsection \ref{sec:conv_method} describes the sufficient conditions of convergence
of the considered method.
The results of the numerical experiments confirming the efficiency of considered method are
presented in Section \ref{sec:exps}.
Conclusion gives a short summary of the results of the study and proposes the
directions for further improvement of the software implementation of the considered method.

\section{Statement of the global optimization problem}
\label{sec:problem}
This paper considers the following problem:
to find the global minimum of an  \(N\)-dimensional function \(\varphi(y)\) in
a hyperinterval $D$
\begin{equation}
\label{eq:task}
\varphi(y^*)=\min\{\varphi(y):y\in D\},
\end{equation}
\begin{displaymath}
D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}.
\end{displaymath}
In order to produce an estimate of the global minimum based on a finite number of
computations of the function values, the rate of variation of \(\varphi(y)\) in \(D\) has to
be limited.
As a rule, the Lipschitz condition
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty,
\end{displaymath}
is accepted as such a limitation.

There are various methods for solving the considered multidimensional problem directly
\cite{SergeyevKvasov2017, Jones2009} as well as efficient methods for solving the univariate problems \cite{Norkin1992, Strongin2000}.
This paper considers a one-dimensional method, which is applied jointly with
the dimensionality reduction scheme.
The use of space¬filling curve (or \textit{evolvent}) $y(x)$, where
\begin{equation}
\label{cube}
\lbrace y\in \mathbf{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\},
\end{equation}
 is a well-known scheme of dimensionality reduction
 of the initial problem for the global optimization algorithms
\cite{Sergeyev2013}.
A mapping of the type (\ref{cube}) allows reducing a multidimensional problem to a univariate one at the expense of worsening its properties. In particular, the
one-dimensional function \(\varphi(y(x))\) is not a Lipschitz function but a Holder one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}}, \; x_1,x_2\in[0,1],
\end{displaymath}
where the H\"{o}lder constant \(H\) is related to the Lipschitz constant \(L\) as
\begin{displaymath}
  H=2L\sqrt{N+3}.
\end{displaymath}

@@

Let us set \(g_{m+1}(y)=\varphi(y)\). Hereafter, we shall assume all functions
\(g_k(y),1\leqslant k \leqslant m+1\),
to satisfy the Lipschitz condition in a certain hyperinterval including \(D\).

@@

Similar to \cite{BarkalovStrongin2018}, this work aims to formulate a method that will ensure a uniform convergence of the
solutions of all problems in the series. Here and below, a uniform convergence implies a proportional decrease
of the distance between the best iteration point and the
global optimizer in all the problems in the series:
\begin{equation}
  \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
    \frac{\Vert \tilde{y_i}(s)^* - y_i*\Vert_\infty}{\Vert \tilde{y_j}(s)^* - y^*_j\Vert_	\infty} \leqslant \varepsilon,
\end{equation}
where \(s\) is the number of steps of the optimization method, \(\tilde{y_i}(s)^*\) is the best iteration point
in the problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).

@@

Taking into account the dimensionality reduction scheme (\ref{cube}), let us assume that the method
requires finding the global minimum of the function \(\varphi(x),
x\in[0,1]\), which satisfies the H\"{o}lder condition with the constraints \(g_j(x)\), which in turn satisfy
this condition in the interval \([0,1]\).

The index algorithm of global search (IAGS) for solving one-dimensional problems
%(\ref{eq:constrained_problem})
considered here implies construction of a sequence of points
\(x_k\), at which the values of index function \(z_k\) are calculated.
The index scheme \cite{Strongin2000} was used to account for the latter.
Let us assume that each function \(g_i(x), 1 \leqslant i \leqslant m + 1\), is defined and computable
only in the corresponding subrange \(Q_i \in [a, b]\), where \(\varphi(x)\) is denoted as \(g_{m+1}(x)\) and
\begin{equation}
  \label{eq:q}
  Q_1=[0,1],\: Q_{i+1}=\left\{x\in Q_i:g_{i}(x)\leqslant 0\right\},\:1 \leqslant i \leqslant m, \: Q_{m+2}=\emptyset
\end{equation}
Considering the definition (\ref{eq:q}) the initial problem can be rewritten as
\begin{displaymath}
  \varphi(x^*)=\min\{g_{m+1}(x):x\in Q_{m+1}\}.
\end{displaymath}

The index scheme assigns a number called index \(\nu(x)\) to each point of the search sequence \(x_k\):
\begin{equation}
  \nu(x) = i:x\in Q_i, x\not\in Q_{i+1},\:1\leqslant i\leqslant m + 1.
\end{equation}

In order to obtain the index of the point \(x\) we have to perform a trial defined via the following steps:
\begin{enumerate}
  \item Sequentially  compute functions \(g_i(x),1 \leqslant i \leqslant m\) until for some \(i^*\) \(g_{i^*}(x)>0\) or until \(i=m+1\).
  Set \(\nu(x)=i^*\) or \(\nu(x)=m+1\) if \(g_i(x) \leqslant 0, 1 \leqslant i \leqslant m\).
  Therefore, \(\nu(x)\) is the number of the first violated constraint at the point \(x\) or \(m+1\) if
  all the constraints are satisfied at \(x\).
  \item Return pair \(\nu(x), z=g_{\nu(x)}(x)\) as a result of the trial.
\end{enumerate}

This approach to trials allows reducing the initial problem with functional
constraints to an unconstrained problem of minimization of a discontinuous function:
\begin{displaymath}
  \begin{array}{lr}
    \psi (x^{*})=\min_{x\in [0,1]}\psi (x), \\
    \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu },&\nu <M,\\(g_{M}(x)-g_{M}^{*})/H_{M},&\nu
=M.\end{cases}}
  \end{array}
\end{displaymath}
Here \(M=\max_{}^{}\left\{\nu (x):x\in [0,1]\right\}\) and \(g_{M}^{*}=\min
_{}^{}\left\{g_{M}(x):x\in Q_{M}\right\}\).
Because of the definition of the number \(M\), the problem of finding \(g_{M}^{*}\) always
has a solution. If \(M=m+1\), then \(g_{M}^{*}=\varphi(x^{*})\).
The function \(\psi (x)\) satisfies the H\"{o}lder condition on the set
\(Q_1\) with the constant 1, and \(\psi (x)\) can have jump discontinuities at the boundaries
of the sets \(Q_i,\:1\leqslant i \leqslant m+1\).
Despite the values of the H\"{o}lder constants \(H_k\) and the value \(g_{M}^{*}\) are not
known in advance, they can be estimated in the course of solving the problem.
A set of triples \(\{(x_k,\nu_k,z_k)\}, 1\leqslant k\leqslant n\), constitutes the search information
accumulated by the method upon execution of \(n\) steps.

At the first iteration of the method, the trial is performed in the arbitrary internal point \(x_1\) of
the interval \([0,1]\).
The indices of the points 0 and 1 are considered to be zero indices, the values \(z\) at these points
are undefined.
Let us assume that \(k\geqslant 1\) iterations of the method have been performed. In the course of this performance the
trials were conducted at \(k\) points \(x_i, 1\leqslant i\leqslant k\).
Then, the points \(x^{k+1}\) of the search trials of the next \((k+1)^{\rm th}\) iteration are
defined according to the rules:

Step 1. Reassign lower indices to the points of the set \(X_k=\{x^1,\dotsc,x^k\}\cup\{0\}\cup\{1\}\), which
includes the boundary points of the interval \([0,1]\) as well as the points of preceding trials in the order of increasing coordinate values, i.e.
\begin{equation}
  \label{eq:points}
  0=x_0<x_1<\dotsc<x_{k+1}=1,
\end{equation}
and compare them with the values \(z_{i}=g_{\nu }(x_{i}),\nu =\nu (x_{i}),i={\overline {1,k}}\) computed at these points.

Step 2. For each integer number \(\nu ,1\leqslant \nu \leqslant m+1\), determine the corresponding
set \(I_{\nu }\) of the lower indices of the points, at which the values of the functions \(g_{\nu
}(x)\) are computed:
\begin{displaymath}
  I_{\nu }=\{i:\nu (x_{i})=\nu ,1\leqslant i\leqslant k\},1\leq \nu \leqslant m+1,
\end{displaymath}
and determine the maximum value of the index \(M=\max\{\nu (x_{i}),1\leq i\leq k\}\).

Step 3. Compute current estimate for the unknown H\"{o}lder constant:
\begin{equation}
  \label{step2}
  \mu _{\nu }=\max \left\{ \frac{|g_{\nu }(x_{i})-g_{\nu }(x_{j})|}{(x_{i}-
x_{j})^{\frac{1}{N}}}:i,j\in I_{\nu },i>j \right\}.
\end{equation}
If the set \(I_{\nu }\) contains less than two elements or if the value \(\mu _{\nu }\) is equal to zero, then assume \(\mu _{\nu }=1\).

Step 4. For all the nonempty sets \(I_{\nu },\nu ={\overline {1,M}}\) compute the estimates
\begin{equation}
  \label{eq:step_4}
  z_{\nu }^{*}={\begin{cases}\min\{g_{\nu }(x_{i}):x_{i}\in I_{\nu }\},&\nu =M,\\
  0,&\nu <M.\end{cases}}
\end{equation}

Step 5. For each interval \((x_{i-1},x_{i}),1\leqslant i\leqslant k\) compute the characteristic
\begin{equation}
  \label{step3_1}
  R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu
})^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}},&\nu =\nu
(x_{i})=\nu (x_{i-1}),\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu
}}},&\nu =\nu (x_{i-1})>\nu (x_{i}),\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu
_{\nu }}},&\nu =\nu (x_{i})>\nu (x_{i-1}),\end{cases}}
\end{equation}
where \(\Delta_{i}=(x_{i}-x_{i-1})^{\frac{1}{N}}\).
The values \(r_{\nu }>1,\nu ={\overline {1,m}}\) are the parameters of the algorithm.
They define the products \(r_{\nu}\mu_{\nu}\) used in computing the characteristics as the estimates of the
unknown H\"{o}lder constants.

@@

The algorithm stops if the condition \(\Delta_{t}\leqslant \varepsilon\) is satisfied. Here
\(\varepsilon>0\) is a predefined precision.
The values
\begin{equation}
\varphi_k^*=\min_{1\leqslant i \leqslant k}\varphi(x_i), \; x_k^*=\arg \min_{1\leqslant i \leqslant
k}\varphi(x_i)
\end{equation}
are assumed as the estimates of the global solution.

Next, following the approach described in \cite{BarkalovStrongin2018}, we shall use \(q\) copies
of IAGS working synchronously to solve the problem series (\ref{eq:many_problems}).
The only difference is that when selecting the interval with the maximum characteristic at Step
6, the choice will be made from all intervals, which were generated by \(q\) copies of IAGS.
If the maximum characteristic corresponds to the \(i^{\rm th}\) problem, then Step 7 is
performed in the \(i^{\rm th}\) copy of the method while the other copies stay idle.
This way, at every iteration the trial is performed for the most promising problem in terms of the characteristics (\ref{step3_1}). This allows distributing the resources of the
method among the problems dynamically. Let's denote this method as MIAGS.

The parallel modification of the method does not differ from the one considered in
\cite{BarkalovStrongin2018} and consists in the selection of \(p\) intervals at Step 6 and
performing  \(p\) trials in parallel at the following step.
All resources of the method within the framework of the iteration may be focused on
a single problem as well as on \(l\leqslant p\) problems simultaneously (depending on the problem,
to which the intervals selected by the method belong).

@@

The conditions of convergence of the method described in Section \ref{sec:method} in case of \(q=1\) are given in \cite{Strongin2000}.
\begin{theorem} (Sufficient convergence conditions)
  \label{th:single_conv}
  Let us assume that the following conditions are true:
  \begin{enumerate}
    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitz functions with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  Then any limit point \(\overline{y}\) of the sequence \(\{y_k\} = \{y(x_k)\}\) generated
  by the index method, the problem
  (\ref{eq:constrained_problem}) is feasible and satisfies the conditions
\begin{equation}
  \label{eq:conv_cond}
  \varphi(\overline{y})=\inf\{ \varphi(y^k): g_i(y^k)\leqslant 0,1\leqslant i\leqslant m, k=1,2,\dots\}=\varphi(y^*).
\end{equation}
\end{theorem}

@@

\subsection{Example of solving a multicriteria problem}

In order to demonstrate the efficiency of the approach in the balancing of the load, let us
consider an example, in which a set of problems of the kind (\ref{eq:many_problems}) is
generated as a result of scalarization of a multicriteria optimization problem with constrains.
